---
title: "Highlighting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Highlighting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(muiCharts)
library(dplyr)
library(tidyr)
```

Highlighting provides visual feedback when users interact with chart elements. This article demonstrates highlighting options based on the [MUI Highlighting documentation](https://mui.com/x/react-charts/highlighting/).

## Highlight Scope

The `highlightScope` prop controls how items are emphasized on hover. It has two properties:

- `highlight` — which items to highlight: `"item"`, `"series"`, or `"none"`
- `fade` — which items to fade: `"global"`, `"series"`, or `"none"`

## Highlight Item

Highlight a single item on hover, fading all others:

```{r}
starwars_gender <- dplyr::starwars |>
  unnest_longer(films) |>
  count(films, sex) |>
  filter(!is.na(sex)) |>
  pivot_wider(names_from = sex, values_from = n) |>
  mutate(release = c(1977, 2002, 1983, 2005, 1980, 2015, 1999)) |>
  arrange(release)

BarChart(
  dataset = starwars_gender,
  xAxis = list(
    list(scaleType = "band", dataKey = "films", tickLabelStyle = list(fontSize = "0.5em"))
  ),
  series = list(
    list(
      dataKey = "male",
      label = "Male",
      highlightScope = list(highlight = "item", fade = "global")
    ),
    list(
      dataKey = "female",
      label = "Female",
      highlightScope = list(highlight = "item", fade = "global")
    )
  ),
  height = 300
)
```

## Highlight Series

Highlight an entire series when hovering over any item in it:

```{r}
BarChart(
  dataset = starwars_gender,
  xAxis = list(
    list(scaleType = "band", dataKey = "films", tickLabelStyle = list(fontSize = "0.5em"))
  ),
  series = list(
    list(
      dataKey = "male",
      label = "Male",
      highlightScope = list(highlight = "series", fade = "global")
    ),
    list(
      dataKey = "female",
      label = "Female",
      highlightScope = list(highlight = "series", fade = "global")
    ),
    list(
      dataKey = "hermaphroditic",
      label = "Hermaphroditic",
      highlightScope = list(highlight = "series", fade = "global")
    )
  ),
  height = 300
)
```

## Faded Item Styling

Customize the appearance of faded items using the `faded` property in pie charts:

```{r}
df <- starwars |>
  count(sex) |>
  filter(!is.na(sex)) |>
  select(label = sex, value = n)

PieChart(
  series = list(
    list(
      data = df,
      highlightScope = list(highlight = "item", fade = "global"),
      faded = list(
        innerRadius = 30,
        additionalRadius = -30,
        color = "gray"
      )
    )
  ),
  height = 300
)
```

## Line Chart Highlighting

Highlighting works with line charts too:

```{r}
LineChart(
  dataset = starwars_gender,
  xAxis = list(
    list(scaleType = "point", dataKey = "films", tickLabelStyle = list(fontSize = "0.5em"))
  ),
  series = list(
    list(
      dataKey = "male",
      label = "Male",
      highlightScope = list(highlight = "series", fade = "global")
    ),
    list(
      dataKey = "female",
      label = "Female",
      highlightScope = list(highlight = "series", fade = "global")
    )
  ),
  height = 300
)
```

## Scatter Chart Highlighting

Scatter charts also support highlighting by series:

```{r}
df_male <- starwars |>
  filter(sex == "male", !is.na(mass), !is.na(height)) |>
  select(x = mass, y = height)

df_female <- starwars |>
  filter(sex == "female", !is.na(mass), !is.na(height)) |>
  select(x = mass, y = height)

ScatterChart(
  series = list(
    list(
      data = df_male,
      label = "Male",
      highlightScope = list(highlight = "series", fade = "global")
    ),
    list(
      data = df_female,
      label = "Female",
      highlightScope = list(highlight = "series", fade = "global")
    )
  ),
  xAxis = list(list(min = 0, label = "Mass (kg)")),
  yAxis = list(list(min = 0, label = "Height (cm)")),
  height = 350
)
```

## Axis Highlight in Composition

When using composition, add `ChartsAxisHighlight` to display a highlight band or line on hover:

```{r}
ChartDataProvider(
  dataset = starwars_gender,
  height = 300,
  xAxis = list(
    list(scaleType = "band", dataKey = "films", tickLabelStyle = list(fontSize = "0.5em"))
  ),
  series = list(
    list(type = "bar", dataKey = "male", label = "Male"),
    list(type = "bar", dataKey = "female", label = "Female")
  ),
  ChartsTooltip(trigger = "axis"),
  ChartsSurface(
    BarPlot(),
    ChartsAxisHighlight(x = "band"),
    ChartsXAxis(),
    ChartsYAxis()
  )
)
```

## Learn More

For more details about highlighting, visit the [MUI X Highlighting documentation](https://mui.com/x/react-charts/highlighting/).
