---
title: "Introduction"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Install

You can install the development version of muiCharts like so:

``` r
remotes::install_github("lgnbhl/muiCharts")
```

## Quick Start

```{r message=FALSE, warning=FALSE}
library(muiCharts)
library(dplyr)

starwars_summary <- dplyr::starwars |>
  filter(species %in% c("Human", "Droid", "Wookiee")) |>
  group_by(species) |>
  summarise(height = mean(height, na.rm = TRUE), mass = mean(mass, na.rm = TRUE))
```

```{r}
BarChart(
  dataset = starwars_summary,
  xAxis = list(
    list(
      scaleType = "band",
      dataKey = "species",
      label = "Species"
    )
  ),
  series = list(
    list(dataKey = "height", label = "Height (cm)"),
    list(dataKey = "mass", label = "Mass (kg)")
  ),
  grid = list(horizontal = TRUE),
  height = 300
)
```

## Using JavaScript

Use the `JS()` function to pass JavaScript code, for example to customize tooltip values with `valueFormatter`:

```{r}
customValueFormatter <- function(category){
  JS(
    paste0(
      "function(value){",
      "if (value === null) { return 'None' }",
      "return `${value} ", category, "`}"
    )
  )
}

BarChart(
  dataset = starwars_summary,
  xAxis = list(
    list(
      scaleType = "band",
      dataKey = "species",
      label = "Species",
      valueFormatter = JS("(value) => `${value}`")
    )
  ),
  series = list(
    list(
      dataKey = "height",
      label = "Height (cm)",
      valueFormatter = customValueFormatter("cm")
    ),
    list(
      dataKey = "mass",
      label = "Mass (kg)",
      valueFormatter = customValueFormatter("kg")
    )
  ),
  grid = list(horizontal = TRUE),
  height = 400
)
```

## Chart Types

muiCharts provides the following chart types, each with a dedicated article:

- [Bar Charts](bars.html) — vertical and horizontal bar charts with grouping and stacking
- [Line Charts](lines.html) — line charts with marks, curves, and null handling
- [Area Charts](areas.html) — filled area charts with stacking and baselines
- [Pie Charts](pies.html) — pie and donut charts with arc labels
- [Scatter Charts](scatter.html) — scatter plots with Voronoi interaction
- [Radar Charts](radar.html) — radar charts for multi-metric comparisons
- [Sparkline Charts](sparkline.html) — compact inline charts
- [Gauge Charts](gauge.html) — gauge displays for single metrics

## Features

Cross-cutting features that apply to all chart types:

- [Axis & Grid](axis.html) — scale types, labels, value formatters, grid lines, margins
- [Tooltip](tooltip.html) — tooltip triggers, formatting, and composition
- [Legend](legend.html) — legend position, direction, and color legends
- [Highlighting](highlighting.html) — highlight and fade interactions
- [Styling](styling.html) — colors, color maps, and CSS overrides with `sx`
- [Composition](composition.html) — build charts from individual sub-components

## Integration

- [Shiny](shiny.html) — reactive charts, click events, and dashboards
- [Quarto](quarto.html) — HTML documents, dashboards, and tabsets
- [Error Bars](ErrorBar.html) — custom error bar components

### Contribute

If you have any issue, question or want to contribute with a pull request, don't hesitate to write me on https://github.com/lgnbhl/muiCharts

For updates follow [Felix Luginbuhl](https://linkedin.com/in/FelixLuginbuhl) on LinkedIn.
