---
title: "Introduction"
format:
  html:
    minimal: true
    code-fold: true
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Install

You can install the development version of muiCharts like so:

``` r
remotes::install_github("lgnbhl/muiCharts")
```

## Quick Start

```{r message=FALSE, warning=FALSE}
library(muiCharts)
library(dplyr)
```

```{r}
BarChart(
  dataset = head(starwars_people, 4),
  xAxis  = list(list(scaleType = "band", dataKey = "name")),
  series = list(list(dataKey = "height", label = "Height (cm)")),
  height = 300
)
```

## Using JavaScript

Use the `JS()` function to pass JavaScript code, for example to customize tooltip values with `valueFormatter`:

```{r}
BarChart(
  dataset = head(starwars_people, 4),
  xAxis  = list(list(scaleType = "band", dataKey = "name")),
  series = list(
    list(dataKey = "height", label = "Height (cm)", valueFormatter = JS("(v) => v + ' cm'")),
    list(dataKey = "mass",   label = "Mass (kg)",   valueFormatter = JS("(v) => v + ' kg'"))
  ),
  height = 300
)
```

## MUI ecosystem

muiCharts integrates seamlessly with [muiMaterial](https://felixluginbuhl.com/muiMaterial) to create modern, fully-styled dashboards.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(),
  {
    primaryColor <- "#1976d2"

    ThemeProvider(
      theme = list(palette = list(primary = list(main = primaryColor))),
      Box(sx = list(p = 3, bgcolor = "background.paper"),
        Card(sx = list(boxShadow = 3),
          CardContent(
            Stack(spacing = 1.5,
              Typography("Character Count", variant = "h6", sx = list(fontWeight = 600)),
              Typography("Number of characters by Star Wars episode", variant = "body2", sx = list(color = "text.secondary")),
              BarChart(
                dataset = starwars_films |> mutate(Characters = lengths(characters)) |> arrange(episode_id),
                xAxis = list(list(scaleType = "band", dataKey = "episode_id")),
                series = list(list(dataKey = "Characters", color = primaryColor)),
                height = 280
              ),
              Divider(),
              Typography("Made with muiCharts and muiMaterial", variant = "caption", sx = list(color = "text.secondary"))
            )
          )
        )
      )
    )
  }
)
```

## Chart Types

muiCharts provides the following chart types, each with a dedicated article:

- [Bar Charts](bars.html) — vertical and horizontal bar charts with grouping and stacking
- [Line Charts](lines.html) — line charts with marks, curves, and null handling
- [Area Charts](areas.html) — filled area charts with stacking and baselines
- [Pie Charts](pies.html) — pie and donut charts with arc labels
- [Scatter Charts](scatter.html) — scatter plots with Voronoi interaction
- [Radar Charts](radar.html) — radar charts for multi-metric comparisons
- [Sparkline Charts](sparkline.html) — compact inline charts
- [Gauge Charts](gauge.html) — gauge displays for single metrics

## Features

Cross-cutting features that apply to all chart types:

- [Axis & Grid](axis.html) — scale types, labels, value formatters, grid lines, margins
- [Tooltip](tooltip.html) — tooltip triggers, formatting, and composition
- [Legend](legend.html) — legend position, direction, and color legends
- [Highlighting](highlighting.html) — highlight and fade interactions
- [Styling](styling.html) — colors, color maps, and CSS overrides with `sx`
- [Composition](composition.html) — build charts from individual sub-components

## Integration

- [Shiny](shiny.html) — reactive charts, click events, and dashboards
- [Quarto](quarto.html) — HTML documents, dashboards, and tabsets

### Contribute

If you have any issue, question or want to contribute with a pull request, don't hesitate to write me on https://github.com/lgnbhl/muiCharts

For updates follow [Felix Luginbuhl](https://linkedin.com/in/FelixLuginbuhl) on LinkedIn.
