---
title: "Radar Charts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Radar Charts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(muiCharts)
library(dplyr)
library(tidyr)
```

This article demonstrates how to create radar charts using muiCharts, based on the [MUI Radar Chart documentation](https://mui.com/x/react-charts/radar/).

## Basics

Create a basic radar chart showing character counts by sex across different metrics:

```{r}
# Prepare data: count characters by sex for each film
starwars_film_sex <- starwars |>
  unnest_longer(films) |>
  count(films, sex) |>
  filter(!is.na(sex)) |>
  pivot_wider(names_from = sex, values_from = n, values_fill = 0)

# Create radar chart for one film
film_data <- starwars_film_sex |>
  filter(films == "A New Hope")

RadarChart(
  height = 300,
  series = list(
    list(
      label = "A New Hope",
      data = c(film_data$male, film_data$female, film_data$hermaphroditic, film_data$none)
    )
  ),
  radar = list(
    metrics = c("male", "female", "hermaphroditic", "none")
  )
)
```

## Multi-series Radar Chart

Compare multiple films on the same radar chart:

```{r}
# Prepare data for three films
film1 <- starwars_film_sex |> filter(films == "A New Hope")
film2 <- starwars_film_sex |> filter(films == "Attack of the Clones")
film3 <- starwars_film_sex |> filter(films == "Return of the Jedi")

RadarChart(
  height = 350,
  series = list(
    list(
      label = "A New Hope",
      data = c(film1$male, film1$female, film1$hermaphroditic, film1$none)
    ),
    list(
      label = "Attack of the Clones",
      data = c(film2$male, film2$female, film2$hermaphroditic, film2$none)
    ),
    list(
      label = "Return of the Jedi",
      data = c(film3$male, film3$female, film3$hermaphroditic, film3$none)
    )
  ),
  radar = list(
    metrics = c("male", "female", "hermaphroditic", "none")
  )
)
```

## Series Options

Customize series appearance with `hideMark` and `fillArea`:

```{r}
RadarChart(
  height = 350,
  series = list(
    list(
      label = "A New Hope",
      data = c(film1$male, film1$female, film1$hermaphroditic, film1$none),
      fillArea = TRUE
    ),
    list(
      label = "Attack of the Clones",
      data = c(film2$male, film2$female, film2$hermaphroditic, film2$none),
      hideMark = TRUE
    )
  ),
  radar = list(
    metrics = c("male", "female", "hermaphroditic", "none")
  )
)
```

## Metrics

Define metrics with detailed configuration using objects with `name`, `min`, and `max`:

```{r}
RadarChart(
  height = 350,
  series = list(
    list(
      label = "A New Hope",
      data = c(film1$male, film1$female, film1$hermaphroditic, film1$none)
    ),
    list(
      label = "Attack of the Clones",
      data = c(film2$male, film2$female, film2$hermaphroditic, film2$none)
    ),
    list(
      label = "Return of the Jedi",
      data = c(film3$male, film3$female, film3$hermaphroditic, film3$none)
    )
  ),
  radar = list(
    metrics = list(
      list(name = "male", min = 0, max = 30),
      list(name = "female", min = 0, max = 30),
      list(name = "hermaphroditic", min = 0, max = 30),
      list(name = "none", min = 0, max = 30)
    )
  )
)
```

## Grid

Customize the radar grid with `shape` and `divisions`:

```{r}
RadarChart(
  height = 350,
  series = list(
    list(
      label = "A New Hope",
      data = c(film1$male, film1$female, film1$hermaphroditic, film1$none)
    ),
    list(
      label = "Attack of the Clones",
      data = c(film2$male, film2$female, film2$hermaphroditic, film2$none)
    ),
    list(
      label = "Return of the Jedi",
      data = c(film3$male, film3$female, film3$hermaphroditic, film3$none)
    )
  ),
  shape = "circular", # or "sharp"
  divisions = 3,
  radar = list(
    metrics = c("male", "female", "hermaphroditic", "none")
  )
)
```

## Highlight

Control highlighting with the `highlight` prop:

```{r}
RadarChart(
  height = 350,
  series = list(
    list(
      label = "A New Hope",
      data = c(film1$male, film1$female, film1$hermaphroditic, film1$none)
    ),
    list(
      label = "Attack of the Clones",
      data = c(film2$male, film2$female, film2$hermaphroditic, film2$none)
    ),
    list(
      label = "Return of the Jedi",
      data = c(film3$male, film3$female, film3$hermaphroditic, film3$none)
    )
  ),
  highlight = "series",
  radar = list(
    metrics = c("male", "female", "hermaphroditic", "none")
  )
)
```

## Tooltip

Customize tooltip behavior with the `trigger` option:

```{r}
RadarChart(
  height = 350,
  series = list(
    list(
      label = "A New Hope",
      data = c(film1$male, film1$female, film1$hermaphroditic, film1$none)
    ),
    list(
      label = "Attack of the Clones",
      data = c(film2$male, film2$female, film2$hermaphroditic, film2$none)
    )
  ),
  radar = list(
    metrics = c("male", "female", "hermaphroditic", "none")
  ),
  slotProps = list(
    tooltip = list(
      trigger = "axis" # or "item"
    )
  )
)
```

## Learn More

- [MUI Radar Chart Documentation](https://mui.com/x/react-charts/radar/)
