---
title: "Scatter Charts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scatter Charts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(muiCharts)
library(dplyr)
```

Scatter charts express the relation between two variables, using points in a surface. They are ideal for visualizing relationships or correlations, identifying clusters, trends, and outliers in datasets.

This article follows the [MUI Scatter Chart documentation](https://mui.com/x/react-charts/scatter/).

## Basics

Scatter chart series should contain a `data` property with an array of objects. Each object requires `x` and `y` properties, with an optional `id` property.

```{r}
# Prepare starwars data with two variables
starwars_scatter <- starwars |>
  filter(!is.na(height), !is.na(mass), mass < 200) |>
  select(name, height, mass, species) |>
  mutate(
    id = row_number(),
    species_group = case_when(
      species == "Human" ~ "Human",
      species == "Droid" ~ "Droid",
      TRUE ~ "Other"
    )
  )

# Create two series based on species group
series_a <- starwars_scatter |>
  filter(species_group == "Human")

series_b <- starwars_scatter |>
  filter(species_group == "Droid")

ScatterChart(
  height = 350,
  series = list(
    list(
      label = "Human",
      data = lapply(1:nrow(series_a), function(i) {
        list(
          id = series_a$id[i],
          x = series_a$height[i],
          y = series_a$mass[i]
        )
      })
    ),
    list(
      label = "Droid",
      data = lapply(1:nrow(series_b), function(i) {
        list(
          id = series_b$id[i],
          x = series_b$height[i],
          y = series_b$mass[i]
        )
      })
    )
  ),
  xAxis = list(
    list(label = "Height (cm)")
  ),
  yAxis = list(
    list(label = "Mass (kg)")
  )
)
```

## Using a Dataset

You can use the `dataset` prop with `datasetKeys` to specify which columns to use for x and y values:

```{r}
ScatterChart(
  height = 350,
  dataset = starwars_scatter,
  series = list(
    list(
      datasetKeys = list(x = "height", y = "mass"),
      label = "Characters"
    )
  ),
  xAxis = list(
    list(label = "Height (cm)")
  ),
  yAxis = list(
    list(label = "Mass (kg)")
  )
)
```

## Interaction

The chart computes Voronoi cells to map pointer position to the closest element. You can define a maximum radius with `voronoiMaxRadius`:

```{r}
ScatterChart(
  height = 350,
  dataset = starwars_scatter,
  series = list(
    list(
      datasetKeys = list(x = "height", y = "mass"),
      label = "Characters"
    )
  ),
  voronoiMaxRadius = 50,
  xAxis = list(
    list(label = "Height (cm)")
  ),
  yAxis = list(
    list(label = "Mass (kg)")
  )
)
```

Set `voronoiMaxRadius = "item"` to trigger interactions only when hovering exactly over an element:

```{r}
ScatterChart(
  height = 350,
  dataset = starwars_scatter,
  series = list(
    list(
      datasetKeys = list(x = "height", y = "mass"),
      label = "Characters"
    )
  ),
  voronoiMaxRadius = "item",
  xAxis = list(
    list(label = "Height (cm)")
  ),
  yAxis = list(
    list(label = "Mass (kg)")
  )
)
```

## Marker Size

Customize the size of points with the `markerSize` prop:

```{r}
ScatterChart(
  height = 350,
  series = list(
    list(
      label = "Human",
      data = lapply(1:nrow(series_a), function(i) {
        list(
          id = series_a$id[i],
          x = series_a$height[i],
          y = series_a$mass[i]
        )
      }),
      markerSize = 10
    ),
    list(
      label = "Droid",
      data = lapply(1:nrow(series_b), function(i) {
        list(
          id = series_b$id[i],
          x = series_b$height[i],
          y = series_b$mass[i]
        )
      }),
      markerSize = 5
    )
  ),
  xAxis = list(
    list(label = "Height (cm)")
  ),
  yAxis = list(
    list(label = "Mass (kg)")
  )
)
```

## Multiple Axes

You can use multiple y-axes for different series:

```{r}
# Prepare data with different scales
starwars_multi <- starwars |>
  filter(!is.na(height), !is.na(mass), !is.na(birth_year), mass < 200) |>
  select(height, mass, birth_year) |>
  mutate(id = row_number())

ScatterChart(
  height = 350,
  dataset = starwars_multi,
  series = list(
    list(
      datasetKeys = list(x = "height", y = "mass"),
      label = "Mass (kg)",
      yAxisId = "leftAxis"
    ),
    list(
      datasetKeys = list(x = "height", y = "birth_year"),
      label = "Birth Year (BBY)",
      yAxisId = "rightAxis"
    )
  ),
  xAxis = list(
    list(label = "Height (cm)")
  ),
  yAxis = list(
    list(id = "leftAxis", label = "Mass (kg)"),
    list(id = "rightAxis", label = "Birth Year (BBY)")
  )
)
```

## Learn More

- [MUI Scatter Chart Documentation](https://mui.com/x/react-charts/scatter/)

See also: [Styling](styling.html) for color maps, [Composition](composition.html) for advanced layout, [Axis & Grid](axis.html) for grid and axis options.
