---
title: "Tooltip"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tooltip}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(muiCharts)
library(dplyr)
library(tidyr)
```

Tooltips provide extra data when hovering over chart elements. This article demonstrates tooltip customization based on the [MUI Tooltip documentation](https://mui.com/x/react-charts/tooltip/).

## Tooltip Triggers

Tooltips can be triggered in two ways:

- `"item"` — displays data about the hovered element (default for scatter and pie charts)
- `"axis"` — displays data about all items at the current axis position (default for bar and line charts)

### Item Trigger

```{r}
starwars_gender <- dplyr::starwars |>
  unnest_longer(films) |>
  count(films, sex) |>
  filter(!is.na(sex)) |>
  pivot_wider(names_from = sex, values_from = n) |>
  mutate(release = c(1977, 2002, 1983, 2005, 1980, 2015, 1999)) |>
  arrange(release)

BarChart(
  dataset = starwars_gender,
  xAxis = list(
    list(scaleType = "band", dataKey = "films", tickLabelStyle = list(fontSize = "0.5em"))
  ),
  series = list(
    list(dataKey = "male", label = "Male"),
    list(dataKey = "female", label = "Female")
  ),
  slotProps = list(
    tooltip = list(trigger = "item")
  ),
  height = 300
)
```

### Axis Trigger

```{r}
BarChart(
  dataset = starwars_gender,
  xAxis = list(
    list(scaleType = "band", dataKey = "films", tickLabelStyle = list(fontSize = "0.5em"))
  ),
  series = list(
    list(dataKey = "male", label = "Male"),
    list(dataKey = "female", label = "Female")
  ),
  slotProps = list(
    tooltip = list(trigger = "axis")
  ),
  height = 300
)
```

## Disabling Tooltip

Set `trigger = "none"` to completely disable the tooltip:

```{r}
BarChart(
  dataset = starwars_gender,
  xAxis = list(
    list(scaleType = "band", dataKey = "films", tickLabelStyle = list(fontSize = "0.5em"))
  ),
  series = list(
    list(dataKey = "male", label = "Male"),
    list(dataKey = "female", label = "Female")
  ),
  slotProps = list(
    tooltip = list(trigger = "none")
  ),
  height = 300
)
```

## Custom Value Formatters

Use `valueFormatter` in series or axis to control what appears in the tooltip:

```{r}
BarChart(
  dataset = starwars_gender,
  xAxis = list(
    list(
      scaleType = "band",
      dataKey = "films",
      tickLabelStyle = list(fontSize = "0.5em"),
      valueFormatter = JS("(value) => `Movie: ${value}`")
    )
  ),
  series = list(
    list(
      dataKey = "male",
      label = "Male",
      valueFormatter = JS("(value) => `${value} characters`")
    ),
    list(
      dataKey = "female",
      label = "Female",
      valueFormatter = JS("(value) => `${value} characters`")
    )
  ),
  height = 300
)
```

## Tooltip in Composition

When using the composition API, add `ChartsTooltip` explicitly inside `ChartDataProvider`:

```{r}
library(muiMaterial)

ChartDataProvider(
  dataset = starwars_gender,
  height = 300,
  xAxis = list(
    list(
      scaleType = "band",
      dataKey = "films",
      tickLabelStyle = list(fontSize = "0.5em")
    )
  ),
  series = list(
    list(type = "bar", dataKey = "male", label = "Male"),
    list(type = "bar", dataKey = "female", label = "Female")
  ),
  ChartsTooltip(trigger = "axis"),
  ChartsSurface(
    BarPlot(),
    ChartsXAxis(),
    ChartsYAxis()
  )
)
```

## Learn More

For more details about tooltip customization, visit the [MUI X Tooltip documentation](https://mui.com/x/react-charts/tooltip/).
